Let There Be Light: \\ Early Observations of the Normal Type Ia Supernova \abc\ \\Show No Evidence for a Dark Period

### 2017-07-24 doc/ms.tex
Include discussion of some of the faillings of the mixing models in the paper.

Similar to our analysis of the \ion{Si}{2}\,$\lambda$6355 line, we 
can measure velocities and pseudo-equivalent widths (pEWs) of 
\ion{C}{2}\,$\lambda\lambda$6580, 7234. We compare the velocity 
evolution of the \ion{C}{2} lines to \ion{Si}{2} in the right panel 
of Figure \ref{fig:velocity_t_exp} and the pEW evolution is
shown in Figure \ref{fig:ew}. These measurements confirm the qualitative analysis from Figure~\ref{fig:carbon}: namely, the pEW of \ion{C}{2}\,$\lambda$6580 is comparable to that
of \ion{Si}{2}\,$\lambda$6355 at $t \approx -16 \; \mathrm{d}$ and 
the pEW of \ion{C}{2} 
decreases until the feature is no longer detectable 
around $t=-10$ days.

@@

The detection of \ion{C}{2} in SNe Ia spectra is relatively rare
as it requires both unburned carbon, which is likely only present in the outermost layers of the ejecta, and non-local thermal equilibrium effects in order to excite the ionized carbon (e.g., \citealt{2007ApJ...654L..53T}). Spectra obtained around or after $T_{B_\mathrm{max}}$ rarely show \ion{C}{2} as the photosphere has receded from the outermost ejecta, while pre-max spectra show evidence for \ion{C}{2} in $\sim$1/4 of all normal SNe Ia (e.g., \citealt{2011ApJ...732...30P,2012MNRAS.425.1917S,2011ApJ...743...27T}), but the signatures are typically weak. While we caution that the sample of normal SNe Ia with spectra taken within a few days of explosion is small, SN\,2013dy is the only other object known to have strong \ion{C}{2} features like \abc\ \citet{2013ApJ...778L..15Z}. As a counterexample, SN\,2011fe only exhibited weak \ion{C}{2} features in its first spectra
\citep{2012ApJ...752L..26P}. Thus, models of \abc\ must explain the anomalously strong \ion{C}{2} absorption observed shortly after explosion.

@@

We have identified several unusual properties in the early observations of \abc, including: (i) a near-linear photometric rise in the days after explosion, (ii) the lack of a dark phase typical of most SNe Ia, and (iii) the presence of strong \ion{C}{2} absorption. While most SNe Ia are powered purely by radioactive decay, the observed radiation shortly after explosion can also include contributions 
from SN shock cooling or the collision of the SN ejecta with a non-degenerate companion. Here we consider those scenarios as a possible explanation for the early behavior of \abc.

@@

Here we consider a model that has not, as of yet, been discussed in the context of \abc. In the context of SN\,2011fe, \citet{2014MNRAS.441..532D} recently examined pulsational delayed-detonation (PDD) models as an explanation for some SNe Ia. Briefly, PDD models differ from ``standard'' delayed detonation (DD) models in that the initial deflagration causes the WD to expand resulting in the release of some unbound material. As the bound material contracts, eventually a subsequent detonation occurs.\footnote{\citet{2014MNRAS.441..532D} note that the deflagration and detonation in their PDD models are artificially triggered.} An important consequence of this sequence for PDD models is that they naturally produce material that avoids burning, unlike DD models that typically leave no unburnt material. This enables the presence of significantly more carbon in the outer layers of the SN ejecta \citep{2014MNRAS.441..532D}.

While examining observations of SN\,2011fe in the context of DD models, \citet{2014MNRAS.441..532D} find that the models are universally too faint and red at early times, $\sim$24-48 hr after explosion. These discrepancies can be somewhat improved if the newly synthesized $^{56}$Ni is artificially mixed throughout the SN ejecta (see \S\ref{sec:Ni_mixing} for further details on the importance of mixing). Instead, \citet{2014MNRAS.441..532D} find that PDD models provide a better match to observations. Briefly, the diffuse material surrounding the WD heats the outer layers of the SN ejecta leading to a steeper more luminous rise, with bluer colors in the few days after explosion. Importantly, the PDD models are nearly indistinguishable from DD models around and post peak.

Qualitatively, the PDD models provide a better match to the observations of \abc\ than DD models. In particular, PDD models provide higher luminosities and faster rise times in the days after explosion, a more rapid evolution towards blue optical colors, and the formation of strong \ion{C}{2} lines that gradually disappear in the $\sim$week after explosion. Quantitatively, there are still some short comings of the models presented in \citet{2014MNRAS.441..532D}. In particular, the power-law index for the $g$-band rise is $\alpha \approx 3$ for each of the PDD models, which is significantly more steep than $\alpha = 0.98$, which we measure for \abc. The optical colors for \abc\ are blue from our first epoch $\sim$1.7 d after explosion and remain approximately constant in the 10 d after explosion, whereas the PDD models in \citet{2014MNRAS.441..532D} either exhibit distinct color evolution (i.e.\ are not constant) or are too red in the first few days after explosion. Nevertheless, the PDD models presented in \citet{2014MNRAS.441..532D} provide several attractive explanations for the unusual features in the early behavior of \abc, and it may be possible that small adjustments to the model, e.g., additional mixing of the radioactive products from the explosion, can better match \abc.

@@

Having ruled out other possibilities we now consider whether the unusual properties of \abc\ can be explained simply by invoking strong mixing in the SN ejecta. As previously mentioned, strong mixing can lead to a short dark phase and \ion{C}{2} absorption. Here we additionally examine if strong mixing can explain the rapid, near-linear rise in the light curve.

In Figure~\ref{fig:Ni56LC} we compare model calculations from \citet{2016ApJ...826...96P} to the observed photometry of \abc. The \citeauthor{2016ApJ...826...96P} models employ a piston driven explosion to explode a single WD progenitor model. As the piston explosion does not result in any nucleosynthesis, the distribution of $^{56}$Ni in the ejecta must be prescribed by hand, which enables their study of the effects of mixing on the resulting SN emission. Each model employs a fixed $0.5\,\sm$ of $^{56}$Ni that has been distributed throughout the ejecta via boxcar averaging (see their Figure~1). The resulting light curves are synthesized using the SuperNova Explosion Code (\texttt{SNEC}; \citealt{2015ApJ...814...63M}), as shown in Figure~\ref{fig:Ni56LC}. Broadly speaking, the results can be summarized as follows: SN with strong mixing exhibit a rapid almost linear rise and quickly develop blue colors, whereas models where the $^{56}$Ni is confined to the innermost layers of the ejecta remain very faint for days after explosion while exhibiting relatively red optical colors during this period. The fast rise and blue color of \abc\ are consistent with the strong mixing models from \citet{2016ApJ...826...96P}, as shown in Figure~\ref{fig:Ni56LC}. We conclude that the early observations of \abc\ can be explained if the $^{56}$Ni synthesized in the explosion is mixed into the outermost layers of the SN ejecta.

@@

>>>>

Similar to our analysis of the \ion{Si}{2}\,$\lambda$6355 line, we 
can measure velocities and pseudo-equivalent widths (pEWs) of 
\ion{C}{2}\,$\lambda\lambda$6580, 7234. We compare the velocity 
evolution of the \ion{C}{2} lines to \ion{Si}{2} in the right panel 
of Figure \ref{fig:velocity_t_exp} and the pEW evolution is
shown in Figure \ref{fig:ew}. These measurements confirm the qualitative analysis from Figure~\ref{fig:carbon}: namely, the pEW of \ion{C}{2}\,$\lambda$6580 is comparable to that
of \ion{Si}{2}\,$\lambda$6355 at $t \approx -16 \; \mathrm{d}$ and 
the pEW of \ion{C}{2} 
decreases until the feature is no longer detectable 
around $t = -10$ d.

@@

The detection of \ion{C}{2} in SNe Ia spectra is relatively rare
as it requires both unburned carbon, which is likely only present in the outermost layers of the ejecta, and non-local thermal equilibrium effects in order to excite the ionized carbon (e.g., \citealt{2007ApJ...654L..53T}). Spectra obtained around or after $T_{B_\mathrm{max}}$ rarely show \ion{C}{2} as the photosphere has receded from the outermost ejecta, while pre-max spectra show evidence for \ion{C}{2} in $\sim$1/4 of all normal SNe Ia (e.g., \citealt{2011ApJ...732...30P,2012MNRAS.425.1917S,2011ApJ...743...27T}), but the signatures are typically weak. While we caution that the sample of normal SNe Ia with spectra taken within a few days of explosion is small, SN\,2013dy is the only other object known to have strong \ion{C}{2} features like \abc\ \citep{2013ApJ...778L..15Z}. As a counterexample, SN\,2011fe only exhibited weak \ion{C}{2} features in its first spectra
\citep{2012ApJ...752L..26P}. Thus, models of \abc\ must explain the anomalously strong \ion{C}{2} absorption observed shortly after explosion.

@@

Relative to the nearby, normal SN\,2011fe, we have identified several unusual properties in the early observations of \abc, including: (i) a near-linear photometric rise in the days after explosion, (ii) the lack of a dark phase typical of most SNe Ia, and (iii) the presence of strong \ion{C}{2} absorption. While most SNe Ia are powered purely by radioactive decay, the observed radiation shortly after explosion can also include contributions 
from SN shock cooling or the collision of the SN ejecta with a non-degenerate companion. Here we consider those scenarios as a possible explanation for the early behavior of \abc.

@@

Here we consider a model that has not, as of yet, been discussed in the context of \abc. To model SN\,2011fe, \citet{2014MNRAS.441..532D} recently examined pulsational delayed-detonation (PDD) models as an explanation for some SNe Ia. Briefly, PDD models differ from ``standard'' delayed detonation (DD) models in that the initial deflagration causes the WD to expand resulting in the release of some unbound material. As the bound material contracts, eventually a subsequent detonation occurs.\footnote{\citet{2014MNRAS.441..532D} note that the deflagration and detonation in their PDD models are artificially triggered.} An important consequence of this sequence for PDD models is that they naturally produce material that avoids burning, unlike DD models that typically leave no unburnt material. This enables the presence of significantly more carbon in the outer layers of the SN ejecta \citep{2014MNRAS.441..532D}.

When comparing observations of SN\,2011fe to their DD models, \citet{2014MNRAS.441..532D} find that the models are universally too faint and red at early times, $\sim$24-48 hr after explosion.  Instead, \citet{2014MNRAS.441..532D} find that PDD models provide a better match to observations. Briefly, the diffuse material surrounding the WD heats the outer layers of the SN ejecta leading to a steeper more luminous rise, with bluer colors in the few days after explosion. Importantly, the PDD models are nearly indistinguishable from DD models around and post peak.

Qualitatively, the PDD models provide a better match to the observations of \abc\ than DD models. In particular, PDD models provide higher luminosities and faster rise times in the days after explosion, a more rapid evolution towards blue optical colors, and the formation of strong \ion{C}{2} lines that gradually disappear in the $\sim$week after explosion. Quantitatively, there are still some short comings of the models presented in \citet{2014MNRAS.441..532D}. In particular, the power-law index for the $g$-band rise is $\alpha \approx 3$ for each of the PDD models, which is significantly more steep than $\alpha = 0.98$, which we measure for \abc. The optical colors for \abc\ are blue from our first epoch $\sim$1.7 d after explosion and remain approximately constant in the 10 d after explosion, whereas the PDD models in \citet{2014MNRAS.441..532D} either exhibit distinct color evolution (i.e.\ are not constant) or are too red in the first few days after explosion. Nevertheless, the PDD models presented in \citet{2014MNRAS.441..532D} provide several attractive explanations for the unusual features in the early behavior of \abc, and it may be possible that small adjustments to the model (e.g., additional mixing of the radioactive material, adjustments in the explosion energy, etc.) can better match \abc.

@@

Having examined other possibilities we now consider whether the unusual properties of \abc\ can be explained simply by invoking strong mixing in the SN ejecta. As previously mentioned, strong mixing can lead to a short dark phase. Here we additionally examine whether strong mixing can explain the rapid, near-linear rise in the light curve.

In Figure~\ref{fig:Ni56LC} we compare model calculations from \citet{2016ApJ...826...96P} to the observed photometry of \abc. The \citeauthor{2016ApJ...826...96P} models employ a piston driven explosion to explode a single WD progenitor model. As the piston explosion does not result in any nucleosynthesis, the distribution of $^{56}$Ni in the ejecta must be prescribed by hand, which enables the study of the effects of mixing on the resulting SN emission. Each model employs a fixed $0.5\,\sm$ of $^{56}$Ni that has been distributed throughout the ejecta via boxcar averaging (see their Figure~1). The resulting light curves are synthesized using the SuperNova Explosion Code (\texttt{SNEC}; \citealt{2015ApJ...814...63M}), as shown in Figure~\ref{fig:Ni56LC}. Broadly speaking, the results can be summarized as follows: SN with strong mixing exhibit a rapid almost linear rise and quickly develop blue colors, whereas models where the $^{56}$Ni is confined to the innermost layers of the ejecta remain very faint for days after explosion while exhibiting relatively red optical colors during this period. The fast rise and blue color of \abc\ are qualitatively consistent with the strong mixing models from \citet{2016ApJ...826...96P}, as shown in Figure~\ref{fig:Ni56LC}. In fact, if we adopt $t_\mathrm{exp} = t_0 - 0.45 \, \mathrm{d}$, we find that the model with the strongest mixing (orange line in Figure~\ref{fig:Ni56LC} provides an excellent match to the $g$-band observations.

While the models from \citet{2016ApJ...826...96P} provide a good match to the optical photometric evolution of \abc, they consistently overpredict the flux in the UV. The models shown in Figure~\ref{fig:Ni56LC} also overpredict the photospheric velocity of \abc\ by $\sim$2--3,000$\; \mathrm{km \; s}^{-1}$. These discrepancies may be accounted for following an improved treatment of line-blanketing and a reduction in the model explosion energy, respectively. Future modeling efforts may improve the match between the simulations and observations for \abc.

Both the PDD models and ejecta-mixing models show discrepancies with some of the early observations of \abc. Nevertheless, we conclude that one, or both, of these scenarios, which feature qualitatively similar predictions, is the most likely explanation for \abc.

We conclude that the early observations of \abc\ can be explained if the $^{56}$Ni synthesized in the explosion is mixed into the outermost layers of the SN ejecta.

@@

\subsection{The Emerging Sample of Young SNe Ia}
